---
title: "data_manip_for_app"
author: "Cory Sauve"
date: "10/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Outline 

## App 1 - Distributions 

#### Libraries 

```{r}
library(tidyverse)
library(here)
```

#### Data 

Original dataframe 
```{r}
lwqa_data <- read_csv(here("2_data", "lwqa_data_clean.csv"),
                      col_types = cols(micro_ug_L = "d", tn_avg = "d"))
```

Cleaning data for all apps 
```{r}
lwqa_data_clean <- lwqa_data %>% 
  # remove random cols 
  select(-X1, -X1_1) %>% 
  # fix tn_avg col 
  mutate(tn_avg = (tn_epi + tn_hypo) / 2)
```

App 1 data 
```{r}
# Select variables for App 1 
app_1_data <- lwqa_data_clean %>% 
  select(sample_id, lake_name, sa_ha, year, date_sampled, max_depth_m, secchi_m,
         ph_epi:tp_hypo, alk_avg:tp_avg, tn_avg:69)

# Gather variables 
app_1_data <- app_1_data %>% 
  gather(key = "parameter", value = "value", 
         -sample_id, -lake_name, -year, -date_sampled)
  
# Split variables into sample_location column

app_1_data1 <- app_1_data %>% 
  filter(parameter %in% c("max_depth_m", "secchi_m", "sa_ha")) %>% 
  mutate(sample_location = NA) %>% 
  select(sample_id, lake_name, year, date_sampled, parameter, sample_location, value)

app_1_data2 <- app_1_data %>% 
  filter(!parameter %in% c("max_depth_m", "secchi_m", "sa_ha")) %>% 
  separate(parameter, c("parameter", "sample_location"))

# Final df 

app_1_data_final <- rbind(app_1_data1, app_1_data2)

# Editing final df 

app_1_data_final <- app_1_data_final %>% 
  mutate(parameter = ifelse(parameter == "alk", "Alkalinity (mg/L)", parameter),
         parameter = ifelse(parameter == "cond", "Conductivity (mg/L)", parameter),
         parameter = ifelse(parameter == "max_depth_m", "Max depth (m)", parameter),
         parameter = ifelse(parameter == "nh3", "Ammonia (mg/L)", parameter),
         parameter = ifelse(parameter == "no3", "Nitrate (mg/L)", parameter),
         parameter = ifelse(parameter == "orgn", "Organic-N (mg/L)", parameter),
         parameter = ifelse(parameter == "ph", "pH", parameter),
         parameter = ifelse(parameter == "sa_ha", "Surface Area (ha)", parameter),
         parameter = ifelse(parameter == "secchi_m", "Secchi depth (m)", parameter),
         parameter = ifelse(parameter == "srp", "Soluble Reactive Phosphorus (mg/L)", parameter),
         parameter = ifelse(parameter == "tn", "Total Nitrogen (mg/L)", parameter),
         parameter = ifelse(parameter == "tp", "Total Phosphorus (mg/L)", parameter),
         sample_location = ifelse(sample_location == "epi", "Epilimnion", sample_location),
         sample_location = ifelse(sample_location == "hypo", "Hypolimnion", sample_location),
         sample_location = ifelse(sample_location == "avg", "Average", sample_location),
         sample_location = ifelse(is.na(sample_location), "Epilimnion", sample_location)
  )
```

```{r, eval=FALSE}
write.csv(app_1_data_final, file = "app_1_data.csv")
```
  
## App 2 - Lake Type Aggregation 

```{r}

# Select variables for App 1 
app_2_data <- lwqa_data_clean %>% 
  select(sample_id, lake_name, lake_type, sa_ha, year, date_sampled, max_depth_m, secchi_m,
         ph_epi:tp_hypo, alk_avg:tp_avg, tn_avg:69)

# Gather variables 
app_2_data <- app_2_data %>% 
  gather(key = "parameter", value = "value", 
         -sample_id, -lake_name, -year, -date_sampled, -lake_type)
  
# Split variables into sample_location column

app_2_data1 <- app_2_data %>% 
  filter(parameter %in% c("max_depth_m", "secchi_m", "sa_ha")) %>% 
  mutate(sample_location = NA) %>% 
  select(sample_id, lake_name, lake_type, year, date_sampled, parameter, sample_location, value)

app_2_data2 <- app_2_data %>% 
  filter(!parameter %in% c("max_depth_m", "secchi_m", "sa_ha")) %>% 
  separate(parameter, c("parameter", "sample_location"))

# Final df 

app_2_data_final <- rbind(app_2_data1, app_2_data2)

# Editing final df 

app_2_data_final <- app_2_data_final %>% 
  mutate(parameter = ifelse(parameter == "alk", "Alkalinity (mg/L)", parameter),
         parameter = ifelse(parameter == "cond", "Conductivity (mg/L)", parameter),
         parameter = ifelse(parameter == "max_depth_m", "Max depth (m)", parameter),
         parameter = ifelse(parameter == "nh3", "Ammonia (mg/L)", parameter),
         parameter = ifelse(parameter == "no3", "Nitrate (mg/L)", parameter),
         parameter = ifelse(parameter == "orgn", "Organic-N (mg/L)", parameter),
         parameter = ifelse(parameter == "ph", "pH", parameter),
         parameter = ifelse(parameter == "sa_ha", "Surface Area (ha)", parameter),
         parameter = ifelse(parameter == "secchi_m", "Secchi depth (m)", parameter),
         parameter = ifelse(parameter == "srp", "Soluble Reactive Phosphorus (mg/L)", parameter),
         parameter = ifelse(parameter == "tn", "Total Nitrogen (mg/L)", parameter),
         parameter = ifelse(parameter == "tp", "Total Phosphorus (mg/L)", parameter),
         sample_location = ifelse(sample_location == "epi", "Epilimnion", sample_location),
         sample_location = ifelse(sample_location == "hypo", "Hypolimnion", sample_location),
         sample_location = ifelse(sample_location == "avg", "Average", sample_location),
         sample_location = ifelse(is.na(sample_location), "Average", sample_location),
         lake_type = ifelse(lake_type == "natural lake", "Natural Lake", lake_type),
         lake_type = ifelse(lake_type == "impoundment", "Impoundment", lake_type)
  ) %>% 
  filter(sample_location == "Average")

#write.csv(app_2_data_final, "app_2_data.csv")
```

## App 3 - Ecoregion 

```{r}
# Select variables for App 1 
app_3_data <- lwqa_data_clean %>% 
  select(sample_id, lake_name, ecoregion_3, sa_ha, year, date_sampled, max_depth_m, secchi_m,
         ph_epi:tp_hypo, alk_avg:tp_avg, tn_avg:69)

# Gather variables 
app_3_data <- app_3_data %>% 
  gather(key = "parameter", value = "value", 
         -sample_id, -lake_name, -year, -date_sampled, -ecoregion_3)
  
# Split variables into sample_location column

app_3_data1 <- app_3_data %>% 
  filter(parameter %in% c("max_depth_m", "secchi_m", "sa_ha")) %>% 
  mutate(sample_location = NA) %>% 
  select(sample_id, lake_name, ecoregion_3, year, date_sampled, parameter, sample_location, value)

app_3_data2 <- app_3_data %>% 
  filter(!parameter %in% c("max_depth_m", "secchi_m", "sa_ha")) %>% 
  separate(parameter, c("parameter", "sample_location"))

# Final df 

app_3_data_final <- rbind(app_3_data1, app_3_data2)

# Editing final df 

app_3_data_final <- app_3_data_final %>% 
  mutate(parameter = ifelse(parameter == "alk", "Alkalinity (mg/L)", parameter),
         parameter = ifelse(parameter == "cond", "Conductivity (mg/L)", parameter),
         parameter = ifelse(parameter == "max_depth_m", "Max depth (m)", parameter),
         parameter = ifelse(parameter == "nh3", "Ammonia (mg/L)", parameter),
         parameter = ifelse(parameter == "no3", "Nitrate (mg/L)", parameter),
         parameter = ifelse(parameter == "orgn", "Organic-N (mg/L)", parameter),
         parameter = ifelse(parameter == "ph", "pH", parameter),
         parameter = ifelse(parameter == "sa_ha", "Surface Area (ha)", parameter),
         parameter = ifelse(parameter == "secchi_m", "Secchi depth (m)", parameter),
         parameter = ifelse(parameter == "srp", "Soluble Reactive Phosphorus (mg/L)", parameter),
         parameter = ifelse(parameter == "tn", "Total Nitrogen (mg/L)", parameter),
         parameter = ifelse(parameter == "tp", "Total Phosphorus (mg/L)", parameter),
         sample_location = ifelse(sample_location == "epi", "Epilimnion", sample_location),
         sample_location = ifelse(sample_location == "hypo", "Hypolimnion", sample_location),
         sample_location = ifelse(sample_location == "avg", "Average", sample_location),
         sample_location = ifelse(is.na(sample_location), "Average", sample_location),
         ecoregion_3 = ifelse(ecoregion_3 == 57, "Huron-Erie Lake Plain", ecoregion_3),
         ecoregion_3 = ifelse(ecoregion_3 == 56, "MI-IN Till Plain", ecoregion_3),
         ecoregion_3 = ifelse(ecoregion_3 == 54, "Central Corn Belt Plain", ecoregion_3),
         ecoregion_3 = ifelse(ecoregion_3 == 55, "Eastern Corn Belt Plain", ecoregion_3),
         ecoregion_3 = ifelse(ecoregion_3 == 72, "Interior River Lowland", ecoregion_3),
         ecoregion_3 = ifelse(ecoregion_3 == 71, "Interior River", ecoregion_3)
  ) %>% 
  filter(sample_location == "Average")

#write.csv(app_3_data_final, "app_3_data.csv")
```

## App 4 - Plankton 
```{r}

# Select variables for App 1 
app_4_data <- lwqa_data_clean %>% 
  select(sample_id, lake_name, lake_type, ecoregion_3, year, date_sampled, secchi_m, chla_ug_L, pktn_cells_mL:zoo_cells_mL)

# Gather variables 
app_4_data <- app_4_data %>% 
  gather(key = "parameter", value = "value", 
         -sample_id, -lake_name, -lake_type, -ecoregion_3, -year, -date_sampled, -ecoregion_3)

# Editing final df 
app_4_data_final <- app_4_data %>% 
  mutate(parameter = ifelse(parameter == "bg_cell_mL", "Blue-green Density (cells/mL)", parameter),
         parameter = ifelse(parameter == "bg_dom_cell_per", "Blue-green Dominance (% cells)", parameter),
         parameter = ifelse(parameter == "bg_dom_nu_per", "Blue-green Dominance (% NU)", parameter),
         parameter = ifelse(parameter == "bg_num_L", "Blue-green Density (NU/L)", parameter),
         parameter = ifelse(parameter == "chla_ug_L", "Chlorophyll-a (ug/L)", parameter),
         parameter = ifelse(parameter == "diatoms_cells_mL", "Diatom Density (cells/mL)", parameter),
         parameter = ifelse(parameter == "diatoms_nu_L", "Diatoms (NU/L)", parameter),
         parameter = ifelse(parameter == "green_cell_mL", "Green Density (cells/mL)", parameter),
         parameter = ifelse(parameter == "green_num_L", "Green Density (NU/L)", parameter),
         parameter = ifelse(parameter == "other_algae_cells_mL", "Other Algae (cells/mL)", parameter),
         parameter = ifelse(parameter == "other_algae_nu_L", "Other Algae (NU/L)", parameter),
         parameter = ifelse(parameter == "pktn_cells_mL", "Total Plankton (cells/mL)", parameter),
         parameter = ifelse(parameter == "pktn_num_L", "Total Plankton (NU/L)", parameter),
         parameter = ifelse(parameter == "rotifer_num_L", "Rofiter Density (NU/L)", parameter),
         parameter = ifelse(parameter == "secchi_m", "Secchi Depth (m)", parameter),
         parameter = ifelse(parameter == "zoo_num_L", "Zooplankton (NU/L)", parameter),
         ecoregion_3 = ifelse(ecoregion_3 == 57, "Huron-Erie Lake Plain", ecoregion_3),
         ecoregion_3 = ifelse(ecoregion_3 == 56, "MI-IN Till Plain", ecoregion_3),
         ecoregion_3 = ifelse(ecoregion_3 == 54, "Central Corn Belt Plain", ecoregion_3),
         ecoregion_3 = ifelse(ecoregion_3 == 55, "Eastern Corn Belt Plain", ecoregion_3),
         ecoregion_3 = ifelse(ecoregion_3 == 72, "Interior River Lowland", ecoregion_3),
         ecoregion_3 = ifelse(ecoregion_3 == 71, "Interior River", ecoregion_3),
         lake_type = ifelse(lake_type == "natural lake", "Natural Lake", lake_type),
         lake_type = ifelse(lake_type == "impoundment", "Impoundment", lake_type)
   )

write.csv(app_4_data_final, "app_4_data.csv")
```

## App 5 - Trophic State Index 


## App 5 - Water Profile Analyzer 


## App 6 - Data Download 

#### Libraries 

#### Data 

#### Functionality 

## App 7 - Dashboard combining Apps 1-6

#### Libraries 

#### Data 

#### Functionality 















```{r}

df <- read_csv("lwqa_data_clean.csv", 
               col_types = cols(tn_avg = "d",
                                micro_ug_L = "d"))

```
```{r}
df1 <- df %>% select(-X1, -X1_1)
```

```{r}
names(df1)
```

Needs:

Break by sample type (epi, hypo, avg) --> make new dataframes 
Rename to just have parameter 
Gather and make new column `sample_type` (Epilimnion, Hypolimnion, Average)
Bind df's together to make df for Shiny app 

App 1 - Parameter Distrubutions 

App 2 - Lake Type/Ecoregion Aggregations 

App 3 - Plankton 

App 4 - Data Download 

App 5 - Water Profile Analyzer 

App 6 - Maybe make a dashboard of all the data 

```{r}
names(df1)
```

# App 1 data (Distributions)
```{r}
df2 <- df1 %>% 
  mutate(tn_avg = (tn_epi + tn_hypo) / 2)

df3 <- df2 %>% 
  select(sample_id, lake_name, 4:5, 9:

```

```{r}
```



```{r}
names(app_1_data)
```

```{r}
test <- 
  gather(app_1_data, key = "sample_id", value = "value", max_depth_m:tn_hypo) %>% 
  select()
```


```{r}
save(app_1_data, file = "app_1_data.RData")
```













